# General tips about R programming

## Manipulating Dates

### Day of year (DOY) to Date

```{r}
# note that R uses a 0 based index for dates only
# this means it starts counting at 0 rather than 1 when working with dates
as.Date(0, origin = "2016-01-01")

# note that R uses a 0 based index
as.Date(1, origin = "2016-01-01")

# figure out the date for day 189
as.Date(189, origin = "2016-01-01")
```


## data.table
[Frequently Asked Questions about data.table](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-faq.html)


## title?

Add leading/ending zeros to a number, it can be done using `formatC` and `sprintf`.

```{r}
num <- c(0, 1, 10, 1.1, 2.34)

formatC(num, width = 4, format = "d", flag = "0")
formatC(num, digits = 3, format = "f", flag = "0")
```


## Progress bar

I copied the some example code from: [R: Monitoring the function progress with a progress bar](https://ryouready.wordpress.com/2009/03/16/r-monitor-function-progress-with-a-progress-bar/)

Text based progress bar:

```{r}
total <- 20
# create progress bar
pb <- txtProgressBar(min = 0, max = total, style = 3)
for (i in 1:total) {
    Sys.sleep(0.1)
    # update progress bar
    setTxtProgressBar(pb, i)
}
close(pb)
```

A GUI progress bar from `tcltk` package:

```{r}
total <- 20
# create progress bar
pb <- tcltk::tkProgressBar(
    title = "progress bar", min = 0,
    max = total, width = 300
)

for (i in 1:total) {
    Sys.sleep(0.1)
    tcltk::setTkProgressBar(pb, i, label = paste(
        round(i / total * 100, 0),
        "% done"
    ))
}
close(pb)
```

A progress bar using the Windows operating system:

```{r}
total <- 20
# create progress bar
pb <- winProgressBar(
    title = "progress bar", 
    label = "Progress", # must provide label here, otherwise it won't show in the update label
    min = 0, max = total, width = 300
)

for (i in 1:total) {
    Sys.sleep(0.1)
    setWinProgressBar(pb, i, label = paste(
        round(i / total * 100, 0),
        "% done"
    ))
}
close(pb)
```
